MJSAI: Designing a Sound, Configurable, Efficient, and Pushdown Monadic Static~Analyzer for JavaScript

### 2014-12-13 cs252r-fa14-writeup/paper.tex
bibtex stuff
Some stuff on implementation

\newtheorem{proposition}[theorem]{Proposition}

@@

We now describe an AAM-style approach to pushdown static analysis for a standard CEK machine.
%
Recall that a CEK machine is a state machine with three components:
%
\begin{enumerate}
\item The control expression $(c : Exp)$
\item The execution environment $(\rho : Var \rightarrow Val)$
\item The continuation $(k : List Frame)$
\end{enumerate}
%
In AAM-style, we first design an instrument concrete semantics.
%
This instrumented concrete semantics will yield a semantically equivalent
concrete interpreter, but with implementation choices which, after abstraction,
will yield an analysis with different properties.
%
To instrument the semantics for a pushdown analysis we:
\begin{enumerate}
\item Break the recursion between environments and values a. la. \citet{aam}
\item Break the recursion between cons-cells in the list type for continuations.
%
This step is also in the discipline of \citet{aam}, although this particular
choice is not made in their original paper.
%
\item Add a memo-table which caches the results of the analysis.
%
This memo-table caches results \emph{invariant to the continuation present during evaluation}.
\end{enumerate}

From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
\item The control expression $(c : Exp)$ as before
\item The environment $(\rho : Var \rightarrow Addr)$
\item The store $(\sigma : Addr \rightarrow Val)$
\item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
\item The continuation pointer $(\kappa-ptr : KAddr)$

@@

\begin{equation*}
\begin{array}{c c l c l}
  c            &:& Exp        && \\
  \rho         &:& Env        &=& Var   \partto Addr \\
  \sigma       &:& Store      &=& Addr  \partto Value \\
  \kappa\sigma &:& KStore     &=& KAddr \partto Frame \times KAddr \\
  \kappa       &:& KAddr      && \\
  \chi         &:& Cache      &=& Exp \times Store \times Env \partto Value
\end{array}
\end{equation*}

@@

  \node[main node] (101) at (0, 1)
    {$\cek{c}{\rho}{\kappa}$};
  \node[main node] (102) [right of = 101]
    {$\cesk{c}{\rho}{\sigma}{\kappa}$};
  \node[main node] (103) [right of = 102]
    {$\cessk{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}$};
  \node[main node] (104) [right of = 103]
    {$\cesskc{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}{\chi}$};

@@

We reformulated the \lambdajs{} semantics into a reduction machine
semantics. The \lambdajs{} evaluation contexts became our \emph{Frame} data
type. Additionally, we do not divide the evaluation contexts into two sets and
use syntactic stepping rules to implement exceptional control flow. We choose to
implement exceptional control flow with meta functions that crawl the stack
looking for appropriate return points.

@@

>>>>

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}

@@

We now describe an AAM-style approach to pushdown static analysis for a standard CEK machine.
%
Recall that a CEK machine is a state machine with three components:
%
\begin{enumerate}
\item The control expression $(c : Exp)$
\item The execution environment $(\rho : Var \rightarrow Val)$
\item The continuation $(k : List Frame)$
\end{enumerate}
%
In AAM-style, we first design an instrument concrete semantics.
%
This instrumented concrete semantics will yield a semantically equivalent
concrete interpreter, but with implementation choices which, after abstraction,
will yield an analysis with different properties.
%
To instrument the semantics for a pushdown analysis we:
\begin{enumerate}
\item Break the recursion between environments and values a. la. AAM \cite{aam}
\item Break the recursion between cons-cells in the list type for continuations.
%
This step is also in the discipline of AAM \cite{aam}, although this particular
choice is not made in their original paper.
%
\item Add a memo-table which caches the results of the analysis.
%
This memo-table caches results \emph{invariant to the continuation present during evaluation}.
\end{enumerate}

From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
\begin{enumerate}
\item The control expression $(c : Exp)$ as before
\item The environment $(\rho : Var \rightarrow Addr)$
\item The store $(\sigma : Addr \rightarrow Val)$
\item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
\item The continuation pointer $(\kappa-ptr : KAddr)$
\end{enumerate}

@@

\begin{equation*}
\begin{array}{c c l c l}
  c            &:& Exp        && \\
  \rho         &:& Env        &=& Var   \partto Addr \\
  \sigma       &:& Store      &=& Addr  \partto Value \\
  \kappa\sigma &:& KStore     &=& KAddr \partto Frame \times KAddr \\
  \kappa       &:& KAddr      && \\
  \$         &:& Cache      &=& Exp \times Store \times Env \partto Value
\end{array}
\end{equation*}

@@

  \node[main node] (101) at (0, 1)
    {$\cek{c}{\rho}{\kappa}$};
  \node[main node] (102) [right of = 101]
    {$\cesk{c}{\rho}{\sigma}{\kappa}$};
  \node[main node] (103) [right of = 102]
    {$\cessk{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}$};
  \node[main node] (104) [right of = 103]
    {$\cesskc{c}{\rho}{\sigma}{\kappa\sigma}{\kappa}{\$}$};

@@

We reformulated the \lambdajs{} semantics into a state machine with explicit
continuations. The \lambdajs{} evaluation contexts became our \emph{Frame} data
type. Additionally, we do not divide the evaluation contexts into two sets and
use syntactic stepping rules to implement exceptional control flow. We choose to
implement exceptional control flow with meta functions that crawl the stack
looking for appropriate return points.

@@

Our static analysis system is the composition of significant existing work. This
exemplifies the ethos of \maam{}: implement only the semantics and derive
analyses for free. In particular, we have four primary components:

\begin{itemize}
\item The \lambdajs{} Parser \& Desugarer
\item The \lambdajs{} Standard Libarary
\item A \lambdajs{} State Machine with Explicit Continuations
\item The \maam{} Framework
\end{itemize}

The third item is the main contribution of this project.

\subsection{MAAM}

The \maam{} framework separates the semantics of a programming language from the
control flow. This same implementation of \lambdajs{} can be used with different
monads to produce path and flow sensitive analysis, as well as other kinds of
analysis. With an appropriate set of monads, traditional points-to constraints
could be generated.

\subsection{\lambdajs{} State Machine with Explicit Continuations}

We translated the deterministic, reduction semantics of \lambdajs{} directly
into a monadic state machine with explicit continuations. The reduction
semantics maps naturally to a state machine with explicit continuations. The
evaluation contexts become \emph{frames} and nested reduction contexts become a
list of frames.

Adapting the deterministic semantics to a monadic semantics consists primarily
of identifying control branching locations and inserting monadic splits at those
points. We do not insert monadic splits at value joins as this would result in
state space explosions. Inserting monadic splits only at control branches
enables the monad to control path and flow sensitivity.

We replaced substitution in the \lambdajs{} semantics with an
environment. Finally, we replaced the syntactic exception handling code with a
meta-function that searches the stack for the appropriate handler.

We, the authors, found the entire translation process completely straightforward
and mechanical, albeit time consuming due to the size of \lambdajs{} when
including the primitive operators.
